---
title: "Lab 3 Question 2 Bioinformatics"
author: "Duc Doung, Martin Smelik, Raymond Sseguya"
date: "2018 M12 8"
output: pdf_document
---

```{r, library, show=FALSE, include=FALSE}
library(ade4)
library(tidyverse)
library(ape)
library(mvMORPH)
library(ouch)
library(mvSLOUCH)

```


# Assignment 2

## 2.1
```{r, eval=TRUE, show=FALSE, include=FALSE, echo=FALSE}
library(ade4)
data(carni70)
View(carni70)
str(carni70)
names(carni70)

carni70_1 <- carni70[[1]]
carni70_2 <- carni70[[2]]

```
**We analysed the carnivores dataset, data(carni70), from the ade4 R package. The data is about 70 species.**

```{r}
head(carni70$tab, 10)
```


```{r, echo=FALSE}
hist(carni70_2$size, xlab="size trait", main = "Histogram of Size")
boxplot(carni70_2$size, main="Size")

```
**The distribution of the size trait is skewed to the left. In other words, the majority of the 70 species have small sizes. Some of the species have outlier sizes.**


```{r, echo=FALSE}
hist(carni70_2$range, xlab="range trait", main = "Histogram of Range")
boxplot(carni70_2$range, main="Range")

```
**The distribution of the range trait is also generally more balanced however with eveident skewness to the left. There are fewer and fewer species with longer ranges.** 


```{r, echo=FALSE}
ggplot()+aes(x=names(carni70_2)[1], y=scale(carni70_2$size))+
  geom_boxplot()+
  geom_boxplot(aes(x=names(carni70_2[2]), y=scale(carni70_2$range)))+
  labs(y="scaled distribution", x="traits",
       title=" Size and Range compared on the same scale")
### Ursus_arctos has the biggest size
```
**After scaling the data and comparing the distributions of size and range traits on the same scale with boxplots, it is clear that the size trait is more skewed towards low values than the range trait.**


```{r}
rownames(carni70_2[which.max(carni70_2[,"size"]), ])
rownames(carni70_2[which.min(carni70_2[,"size"]), ])
rownames(carni70_2[which.max(carni70_2[,"range"]), ])
rownames(carni70_2[which.min(carni70_2[,"range"]), ])

```
**The species "Ursus_arctos" has the biggest size while "Mustela_nivalis" the smallest size. The species "Puma_concolor" has the biggest range while the species "Bassariscus_pauli" has the smallest range**


```{r}
tree_phylo <- ape::read.tree(text=carni70_1)
plot(tree_phylo)

```
**Above is a plot of the phylogenetic tree.**



## 2.2

### 2.2.1 Both traits evolve as independent Brownian motions
```{r, echo=FALSE}
mvMORPH::mvBM(tree_phylo, carni70$tab, model="BM1", 
             param = list(constraint = "diagonal"))
```
**Because of the assumption that the two traits follow independent processes, the off diagonal elements of the estimated rate matrix are zero. From the rate matrix, we see that the changes in the size trait along the phylogenetic tree are more frequent (at 73) than the changes of the range trait on the same tree. That explains why the histogram of the size trait is more left skewed than that of the range trait (13). This model also suggests that the ancestor(root) had a size of 38.43947 and a range of 13.78439.**


### 2.2.2 The traits evolve as a correlated Brownian motion
```{r, echo=FALSE}
library(ouch)
tree_ouch <- ouch::ape2ouch(tree_phylo, branch.lengths = tree_phylo$edge.length)
library(mvSLOUCH)
mvSLOUCH::BrownianMotionModel(tree_ouch, data = as.matrix(carni70$tab))
```
**Under the assumption that the two traits follow correlated Brownian motion processes, the ancestor (root) had a size of 8.434762 and a range of 11.084473. Even under this model, the size trait changes more frequently than the range trait, with 29383.3025 for size versus 678.3047 for range. The difference between size and range is clearly indicated in the skewnesses of histograms.**



### 2.2.3 independent Ornstein Uhlenbeck processes
```{r}
mvOU(tree_phylo, data=carni70$tab$size, model = c("OU1"), 
     diagnostic = FALSE, echo = TRUE)
mvOU(tree_phylo, data=carni70$tab$range, model = c("OU1"), 
     diagnostic = FALSE, echo = TRUE)
```



### 2.2.4 traits evolve as a bivariate Ornstein-Uhlenbeck process
```{r}
mvOU(tree_phylo, data=carni70$tab, model = c("OU1"), 
     diagnostic = TRUE, echo = TRUE)
```



